---
title: 'p8105_hw2_jc6693'
author: 'Jingfeng Cao'
output: github_document
---

## Problem 1

```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(readxl)
library(haven)
```

import data pols_month
```{r message=FALSE, warning=FALSE}
pols_month_df = read_csv('fivethirtyeight_datasets/pols-month.csv') |>
  janitor::clean_names() |>
  separate(mon,into = c('year','month','day'),sep = '-') |>
  mutate(year = as.integer(year), month = as.integer(month), day = as.integer(day)) |>
  mutate(month = lubridate::month(month,label = TRUE, abbr = FALSE)) |>
  mutate(president = case_when(prez_dem == 1 ~ 'dem', prez_gop == 1 ~ 'gop')) |>
  select(-prez_dem, -prez_gop, -day)
pols_month_df
```

import data snp
```{r message=FALSE, warning=FALSE}
snp_df = read_csv('fivethirtyeight_datasets/snp.csv') |>
  janitor::clean_names() |>
  separate(date,into = c('month','day','year'),sep = '/') |>
  mutate(year = as.integer(year), month = as.integer(month), day = as.integer(day)) |>
  mutate(year = case_when(year <= 25 ~ year+2000, year >= 26 ~ year+1900)) |>
  arrange(year,month) |>
  mutate(month = lubridate::month(month,label = TRUE, abbr = FALSE)) |>
  select(year,month, everything(), -day)
snp_df
```

import data unemployment
```{r message=FALSE, warning=FALSE}
unemployment_df = read_csv('fivethirtyeight_datasets/unemployment.csv') |>
  pivot_longer(Jan:Dec,names_to = 'month',values_to = 'unemployment') |>
  mutate(month = month.name[match(month, month.abb)]) |>
  janitor::clean_names() |>
  mutate(year = as.integer(year))
unemployment_df
```

join and brief description

```{r}
fivethirtyeight_df = left_join(pols_month_df,snp_df,by = c('year','month')) |>
  left_join(unemployment_df,by = c('year','month'))
fivethirtyeight_df
```

the pols_month dataset contains 822 lines and 9 columns related to the number of national politicians who are democratic or republican. snp dataset contains contains 787 lines and 3 columns related to the closing value of Standard & Poor’s stock market index (S&P). the umenployment dataset contains 816 lines and 3 columns related to the percentage of unemployment.   
After merging, the result dataset contains 822 lines and 11 columns. the year ranges from 1947 to 2015, with some missing data of unemployment in 1947 and snp in 1947 to 1949. the key variables are year and month showing time; govgop:president(total 7 variables) showing the politicians; snp showing the stock's values; unemployment showing the unemployment rate.

## Problem 2

import clean the Mr. Trash Wheel sheet
```{r message=FALSE, warning=FALSE}
mr_trash_wheel_df = read_excel('202509 Trash Wheel Collection Data.xlsx', sheet = 1, skip = 1) |>
  select(-...15,-...16) |>
  janitor::clean_names() |>
  head(-1) |>
  mutate(across(everything(), ~replace_na(.x, 0))) |>
  mutate(sports_balls = as.integer(round(sports_balls)))
mr_trash_wheel_df
```

import Professor Trash Wheel and Gwynnda data:
```{r message=FALSE, warning=FALSE}
professor_df = read_excel('202509 Trash Wheel Collection Data.xlsx', sheet = 2, skip = 1) |>
  janitor::clean_names() |>
  head(-1) |>
  mutate(across(everything(), ~replace_na(.x, 0)))
professor_df
gwynnda_df = read_excel('202509 Trash Wheel Collection Data.xlsx', sheet = 4, skip = 1) |>
  janitor::clean_names() |>
  head(-1) |>
  mutate(across(everything(), ~replace_na(.x, 0)))
gwynnda_df
```

join them:
```{r}
mr_trash_wheel_df = mutate(mr_trash_wheel_df, type = 'mr', year = as.integer(year))
professor_df = mutate(professor_df, type = 'professor', year = as.integer(year))
gwynnda_df = mutate(gwynnda_df, type = 'gwynnda', year = as.integer(year))
merge_df = bind_rows(mr_trash_wheel_df,professor_df,gwynnda_df) |>
  arrange(date) |>
  select(type,date,everything(),-dumpster)
merge_df
```

in total, the dataset has 1188 rows of 14 variables, including type showing the data comes from which wheel; date, month, year showing the time of data collected; weight_tons showing the weight of garbage which unit is ton; volume_cubic_yards showing the volume of garbage which unit is cubic yard; from plastic_bottles to sports_balls(total 7 columns) showing the amount of each type garbage; homes_powered showing the amount of homes powered by the electricity generated from garbage.     
The total weight of trash collected by Professor Trash Wheel is `r sum(professor_df$weight_tons)` tons. The total number of cigarette butts collected by Gwynnda in June of 2022 is `r sum(filter(gwynnda_df,month == 'June',year == 2022)$cigarette_butts)`. 










